---
title: "Assessing model agreement in wheat grain nitrogen content prediction - a use case"
author: "Leo Bastos & Adrian Correndo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Assessing model agreement in wheat grain nitrogen content prediction - a use case}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 1 Introduction  
The *`metrica`* package was developed to visualize and compute the level of agreement between observed ground-truth values and model-derived (e.g., mechanistic or empirical) predictions. 

This package is intended to fit into the following workflow:  

1. a data set containing the observed values is used to train a model  
2. the trained model is used to generate predictions  
3. a data frame containing at least the **observed** and model-**predicted** values is created  
4. *`metrica`* package is used to compute goodness of fit and error metrics based on observed and predicted values  
5. *`metrica`* package is used to visualize model fit and selected fit metrics  

This vignette introduces the functionality of the *`metrica`* package applied to observed and model-predicted values of wheat grain nitrogen (N) content (in grams of N $m^{-2}$).  

## 2 Wheat grain N content  
Let's begin by loading the packages needed.  
```{r libraries, message=F, warning=F}
library(ggplot2)
library(dplyr)
library(metrica)
```

Now we load the `wheat` data set included in the `metrica` package.
```{r load data}
# Load
data(wheat)

# Printing first observations
head(wheat)
```

This data set contains two columns:  

- **pred**: model-predicted wheat grain N content, in g N $m^{-2}$,  
- **obs**: ground-truth observed wheat grain N content, in g N $m^{-2}$  

## 3 Visual assessment of agreement  
### 3.1 Scatterplot of pred vs. obs  
The simplest way to visually assess agreement between observed and predicted values is with a scatterplot.  

We can use the function `scatter_plot()` from the *metrica* package to create a scatterplot.  

The function requires specifying at least:  

- the data frame object name (`data` argument)
- the name of the column containing observed values (`obs` argument)  
- the name of the column containing predicted values (`pred` argument)  

Besides a scatterplot, this function also adds to the plot the **1:1 line** (solid line) and the **linear regression line** (dashed line).  

```{r scatter_plot PO}
scatter_plot(data = wheat, 
             obs = obs, 
             pred = pred)
```

The default behavior of `scatter_plot()` places the `obs` column on the x axis and the `pred` column on the y axis (`orientation = "PO"`). This can be inverted by changing the argument `orientation` to "OP":    
```{r scatter_plot OP}
scatter_plot(data = wheat, 
             obs = obs, 
             pred = pred,
             orientation = "OP")
```


The output of the `scatter_plot()` function is a `ggplot2` object that can be further customized:  
```{r scatter_plot custom}
scatter_plot(data = wheat, 
             obs = obs, 
             pred = pred,
             orientation = "OP")+
  labs(x ="Predicted wheat N content (g N/m2)",
       y = "Observed wheat N content (g N/m2)")+
  theme_dark()
```

## 3.2 Bland-Altman plot  
The Bland-Altman plot is another way of visually assessing observed vs. predicted agreement. It plots the difference between observed and predicted values on the y axis, and the observed values on the x axis:  

```{r bland-altman}
bland_altman_plot(data = wheat,
                  obs = obs, 
                  pred = pred)
```

## 4 Numerical assessment of agreement  
The *metrica* package contains functions for **41 metrics** to assess agreement between observed and predicted values for continuous data (i.e., regression error).  

A list with all the the metrics including their name, definition, details, formula, and function name, please check [here].  

All of the metric functions take the same three arguments as the plotting functions:  

- the data frame object name (`data` argument)
- the name of the column containing observed values (`obs` argument)  
- the name of the column containing predicted values (`pred` argument)  

The user can choose to calculate a single metric, or to calculate all metrics at once.  

To calculate a single metric, the metric function can be called.  
For example, to calculate $R^{2}$, we can use the `R2()` function:  
```{r r2}
R2(data = wheat,
   obs = obs, 
   pred = pred)
```

Similarly, to calculate root mean squared error, we can use the `RMSE()` function:  
```{r rmse}
RMSE(data = wheat, 
     obs = obs, 
     pred = pred)
```

The user can also calculate all 41 metrics at once using the function `metrics_summary()`:  
```{r metrics summary}
metrics_summary(data = wheat,
                obs = obs, 
                pred = pred, 
                type = "regression")
```

If the user wants just specific metrics, within the same function `metrics_summary()`, user can pass a list of desired metrics using the argument "metrics_list" as follows:  
```{r metrics summary list}

my.metrics <- c("R2","MBE", "RMSE", "RSR", "NSE", "KGE", "CCC")

metrics_summary(data = wheat,   
                obs = obs,    
                pred = pred,
                type = "regression",
                metrics_list = my.metrics) 

```


## 5 Visual and numerical assessment combined  
The user can also create a scatter plot that includes not only the **predicted** vs. **observed** points, **1:1 line**, and **regression line**, but also **selected metrics and their values** plus the **SMA regression equation**.  

This is accomplished with the function `scatter_plot()`:  

```{r scatter_plot}
scatter_plot(data = wheat,
             obs = obs, 
             pred = pred)

```

To print the metrics on the `scatter_plot()`, just use print.metrics. Warning: do not forget to specify your 'metrics.list':  

```{r scatter_plot print_metrics}

my.metrica.plot <- scatter_plot(data = wheat,
                                obs = obs, 
                                pred = pred,
                                print_metrics = TRUE, metrics_list = my.metrics)

my.metrica.plot

```
Also, as a ggplot element, outputs are flexible of further edition:  

```{r scatter_plot.edit }

my.metrica.plot +
  # Modify labels
  labs(x = "Observed (days to emergence)", y = "Predicted (days to emergence)")+
  # Modify theme
  theme_light()

my.metrica.plot +
  # Modify labels
  labs(x = "Observed (Mg/ha)", y = "Predicted (Mg/ha)")+
  # Modify theme
  theme_dark()
```


## 6 Exporting  
To export the metrics summary table, the user can simply write it to file with the function `write.csv()`:  

```{r export metrics_summary, eval=F }
metrics_summary(data = wheat,
   obs = obs, 
   pred = pred) %>%
  write.csv("metrics_summary.csv")

```


Similarly, to export a plot, the user can simply write it to file with the function `ggsave()`:  

```{r export plot, eval=F}

ggsave(plot = my.metrica.plot,
       "scatter_metrics.png",
       width = 5,
       height = 5)
```


